---
title: "R Tidyverse Methods for SoGE Geographers"
author: "Sami Rifai"
date: "10/15/2018"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

# **R for geographers**
## An exposition of the [Tidyverse](https://www.tidyverse.org/)

<br>
<br>

### Outline
1. **dplyr** & **tibble**: for manipulating data
2. **readr**: for reading and writing data
3. **lubridate**: for parsing dates
4. **tidyr**: for reshaping data
5. **ggplot2**: for visual exploration
6. **sf & stars**: The next gen of geospatial methods for R & Tidyverse methods

<br>
<br>

## If you don't have the necessary libraries, run the following:
```{r cache=T, eval=FALSE}
install.packages(c("tidyverse","lubridate","sf","stars","raster"))
```

## If you don't have Rstudio installed and don't have time, try [RstudioCloud](https://rstudio.cloud/)

# (Section: 1) Basic plumbing with **dplyr** 
- This is a 'pipe' symbol: %>%
- The keyboard shortcut in Rstudio is "ctrl+shift+m" 
- pipes allow chaining operations to an object

<br>

## non pipe way
- this is the 'nested' way to do: generate a vector of random numbers, log it, and plot the distribution
```{r cache=T}
hist(log(rnorm(1000, mean = 100, sd=3)))
```

<br>

## pipe way 
- Here you can linearly read the chain of operations
```{r cache=T}
#  vector object            -> function -> function
rnorm(1000, mean=100, sd=3) %>% log() %>% hist()
```

**So pipes allow us to arbitrarily long things without nesting or creating copyies of dataframes**

<br>
<br>

---------------------------------------------------------------------------

## Life before and after dplyr 
- excellent tutorial: [https://suzan.rbind.io/categories/tutorial/]

<br>
<br>


# - Tibbles vs dataframes - 
## tibbles are a more clever version of data frames that offer some features (?tibble) but also, are compatible for tidyverse methods
```{r cache=T}
junk_df <- data.frame(x=rnorm(20, mean = 0, sd=10), 
                   y=rnorm(20, mean=10, sd=1))
print(junk_df) # prints the whole df!

#*** NOW LET'S MAKE A BIG TIBBLE***#
junk_tb <- tibble(x=rnorm(100, mean = 0, sd=10), 
                      y=rnorm(100, mean=10, sd=1))
print(junk_tb) # just prints the top 10 rows
```

<br>
<br>

# **Filtering with dplyr**
## Base R approach to filtering data: 
```{r cache=T}
iris[iris[,'Species']=="setosa" & iris[,"Sepal.Length"] > 5.0,]
```

## Filtering with **dplyr**
```{r cache=T}
iris %>% filter(Species=="setosa" & Sepal.Length > 5.0)
```

<br>

# Creating or transforming variables
```{r cache=T}
tmp <- iris %>% 
  mutate(p_wl_ratio = Petal.Width/Petal.Length) %>% 
  mutate(narrow = ifelse(p_wl_ratio < 0.25, TRUE, FALSE))

tmp %>% head()
```
<br>


# Renaming variables 
```{r cache=T}
iris %>% names()

iris %>% 
  rename_all(tolower) %>% # rename cols with lowercase
  head() # shows the top rows
```
<br>


# Sample random rows 
```{r cache=T}
iris %>% 
  sample_n(10)
```
<br>
<br>


# **Summarize groups with a statistic**
```{r cache=T, message=F, warning=F}
library(tidyverse)
iris %>% # count number of observations per species
  group_by(Species) %>% # grouping variable
  summarize(nobs = n()) %>% # count the number of observations
  ungroup() # always ungroup, not strictly necessary, but it will save you much pain in time
```
<br>

# Summarize many variables for a corresponding group
```{r cache=T, message=F, warning=F}
library(tidyverse); 
iris %>% # calc mean of traits per species
  group_by(Species) %>% 
  summarise_all(mean) %>% # quick way to generate statistic for many columns
  ungroup()
```
<br>

```{r cache=T, message=F, warning=F}
iris %>% # calc mean of traits across three species
  group_by(Species) %>% 
  summarise_all(mean) %>% 
  ungroup() %>% 
  select(-Species) %>% # de-select Species
  summarize_all(mean) %>% 
  ungroup()
```


<br>
<br>
<br>

# (Sections: 2 & 3) **readr** & **tidyr** for getting data into a workable shape 
## The **readr** package imports data more quickly and with fewer errors.
## It will also not import characters as factors. _unbelievably important_ 
```{r message=F, cache=T}
library(tidyverse); library(lubridate); 

read.table("data/mei_1950_2018.data",skip = 1, nrows = 68) # base R way

mei_wide <- readr::read_table("data/mei_1950_2018.data", skip = 1, col_names = F, n_max = 68) # slightly easier tidyverse way

names(mei_wide) <- c("year",1:12) # add names to columns

mei_long <- mei_wide %>% # reshape the data frame from 'wide' to 'long' 
  gather(key="month",value="index",-year) %>% # use gather() to assemble key-value pairs
  mutate(date=parse_date_time(paste(year,month,1),"ymd")) # mutate the date

mei_long %>% 
  ggplot(data=., aes(date, index))+geom_line()
```

<br>

# Import and summarize data from the EE format
Zonal statistics from Earth Engine are typically exported from Earth Engine with a system:index column.
This can be used to extract the date. Here we will import a time series from four Colombian cities
from the "Climate Hazareds Infrared with Station data" product. The observations are recorded on the 
pentad (5-days). We will aggregate to the month and plot the time series. 

```{r cache=T, message=F, warning=F}
library(tidyverse); 
library(lubridate); 

# rainfall data exported from Google Earth Engine
tmp <- read_csv("data/CHIRPS_ColombiaCiudades_20000101_20180101.csv")

tmp %>% glimpse()

tmp %>% 
  rename(precip=mean, city=name) %>% # rename column
  mutate(date=parse_date_time(substr(`system:index`,1,8),'ymd')) %>% 
  mutate(year=year(date),   # lubridate functions to extract year and month from POSIXct date object
         month=month(date)) %>% 
  group_by(year,month,city) %>% # what groups will we use to summarize the data?
  summarize(precip_tot = sum(precip, na.rm=T)) %>% 
  ungroup() %>% 
  mutate(date = parse_date_time(paste(year,month,1),'ymd')) %>% # reassemble date
  ggplot(data=., aes(date, precip_tot, color=city))+geom_line()+
  labs(x="",y="Precipitation [mm]")


```

<br>
<br>
<br>


# (5) **ggplot2** for data visulization
# ggPlotting El Niño --------------------------------------------------------
## _"always plot your data"_ 
```{r cache=T, message=F, warning=F}
library(tidyverse);

tmp <- read_csv("data/nino34_1870_2017.csv")

# plot the whole record
tmp %>% 
  #__________________x_____y_______thing to add to plot
  ggplot(data=., aes(date, index))+geom_line()

#plot record by month
tmp %>%
  ggplot(data=., aes(x=date, y=index))+ # what are the axis vars?
  geom_line()+                          # what will go on the plot?
  geom_smooth(method='lm',se=F)+        # add a linear trend line
  facet_wrap(~month)                    # generate the plot for every month

# plot recent ENSO record
tmp %>%
  filter(year>=1990) %>%                # filter for years >= 1990
  ggplot(data=., aes(date, index))+
  geom_line()

```

<br>

## LETS 'smooth' the record with a moving average
```{r cache=T, message=F, warning=F}
library(RcppRoll)
tmp %>% 
  arrange(date) %>%  # sort by the date
  mutate(index_12mo = roll_meanr(index, n=12, fill=NA)) %>%  # running 12 month mean
  filter(year>=1990) %>% 
  ggplot(data=., aes(date, index))+
  geom_line()+
  geom_line(aes(date, index_12mo),col='red',lwd=1.5)
```

<br>

## LET's 'deseasonlize' the record by subtracting the monthly mean
```{r cache=T, message=F, warning=F}
df_norms <- tmp %>%
  group_by(month) %>% 
  summarize(index_u = mean(index, na.rm=T)) %>% 
  ungroup()
tmp2 <- left_join(tmp, df_norms, by=c("month")) # now we join it back together

tmp2 %>% 
  mutate(index_ds = index-index_u) %>% 
  filter(year>=1990) %>% 
  ggplot(data=., aes(date, index))+
  geom_line()+
  geom_line(aes(date, index_ds), col='red') # so that actually didn't make much of a difference
```

<br>
<br>

# ggplot spatial data: La Selva CARBONO plots data ---------------------------------------------
```{r cache=T, message=F,  warning=F}
library(tidyverse); library(lubridate)
carb <- read_csv("data/claros1999_2012fulldataset.csv", skip = 5)
carb %>% glimpse()
carb <- carb %>% 
  rename(year=`*Year`) %>% 
  mutate(date = parse_date_time(Date, '%d-%m-%y')) %>% 
  select(-Date)

carb %>% # bad way
  filter(year==1999) %>% 
  ggplot(data=., aes(X_m, Y_m, color=height_cm))+
  geom_point()

carb %>% # better way
  filter(year==1999 & plot=="A1") %>% 
  ggplot(data=., aes(X_m, Y_m, fill=log10(height_cm)))+
  geom_raster()+
  scale_fill_viridis_c()


```

<br>

```{r cache=T, message=F, warning=F}
library(tidyverse); 
carb %>% # visualize through time
  filter(plot=='A1') %>%
  ggplot(data=., aes(X_m, Y_m, fill=log10(height_cm)))+
  geom_raster()+
  scale_fill_viridis_c()+
  facet_wrap(~year)

```

```{r cache=T, eval=F}
#REALLY VISUALIZE it 
# library(gganimate)
# carb %>% # visualize through time
#   filter(year==1999) %>%
#   ggplot(data=., aes(X_m, Y_m, fill=log10(height_cm)))+
#   geom_raster()+
#   scale_fill_viridis_c()+
#   facet_wrap(~plot)

```

## Or visulize it with a movie using [gganimate](https://github.com/thomasp85/gganimate)
```{r cache=T, eval=F}
library(gganimate)
p <- carb %>% # visualize through time
  ggplot(data=., aes(X_m, Y_m, fill=log10(height_cm), frame=year))+
  geom_raster()+
  coord_equal()+
  scale_fill_viridis_c("Canopy Height [log cm]", option = 'B')+
  facet_wrap(~plot)+
  labs(title='Year: {frame_time}')
gganimate(p, "outputs/carbono_plot_heights.gif")
```

<br>

## Plot distributions ------------------------------------------------------
```{r cache=T, eval=F}
library(tidyverse)

hist(carb$height_cm) # old base-R way to plot histogram
plot(density(carb$height_cm)) # base-R way to plot kernel density

carb %>% glimpse
carb %>% ggplot(data=., aes(x=height_cm))+geom_histogram()
carb %>% 
  filter(near(year,2000,tol = 0.1)) %>% # filtering for numbers can be tricky, use near to specify a filter with a tolerance
  ggplot(data=., aes(x=height_cm))+geom_histogram()+facet_wrap(~plot)
carb %>% 
  filter(near(year,2000,tol = 0.1)) %>% 
  ggplot(data=., aes(x= log1p(height_cm)))+
  geom_histogram(bins = 10)+
  scale_y_continuous(trans="log1p")+
  facet_wrap(~plot)

```

<br>
<br>

## Spatiotemporal example --------------------------------------------------
### Plotting monthly ozone concentrations
```{r cache=T, warning=F, message=F}
library(tidyverse)
nasa               # so it's not a tibble
nasa %>% class     # what is the class of the data? 
nasa %>% glimpse() # examine the types of data in the object

nasa %>%                         # spatiotemporal tbl_cube
  as_tibble() %>%                # convert to tibble for ggplot
  ggplot(data=., aes(long,lat))+ #
  geom_raster(aes(fill=ozone))+  # plot the ozone on a lat-long grid
  coord_equal()+                 # 
  scale_fill_viridis_c() + 
  facet_wrap(~month)

```

<br>

### Plot the mean monthly temperature
```{r cache=T, warning=F, message=F}
library(tidyverse)
nasa %>% 
  as_tibble() %>% 
  group_by(lat,long, month) %>% 
  summarize(u=mean(temperature,na.rm=T)) %>% 
  ggplot(data=., aes(long,lat))+
  geom_raster(aes(fill=u))+
  coord_equal()+
  scale_fill_viridis_c() + 
  facet_wrap(~month)

nasa %>% 
  as_tibble() %>% 
  group_by(lat,long, month) %>% 
  summarize(u=mean(temperature,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(tempC = u - 273.15) %>% 
  ggplot(data=., aes(long,lat))+
  geom_raster(aes(fill=tempC))+
  coord_equal()+
  scale_fill_viridis_c() + 
  facet_wrap(~month)

nasa %>% 
  as_tibble() %>% 
  group_by(lat,long, year,month) %>% 
  # summarize(u=mean(temperature,na.rm=T)) %>% 
  # ungroup() %>% 
  mutate(tempC = temperature - 273.15) %>% 
  mutate(hemi = cut(lat,breaks = c(-Inf,0,Inf),labels = c("SH","NH"))) %>% 
  group_by(hemi,year,month) %>% 
  summarize(u=mean(tempC,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(date=parse_date_time(paste(year,month,1),'ymd')) %>% 
  ggplot(data=., aes(date,u,color=hemi))+
  geom_line()+
  scale_fill_viridis_c() 

```

<br>

## Seals -------------------------------------------------------------------
```{r cache=T, message=F, warning=F}
library(tidyverse)
data("seals")     # load example seals data
seals %>% glimpse # check the data types

seals %>% 
  mutate(distance=sqrt(delta_long**2 + delta_lat**2)) %>% # calc the distance travelled
  ggplot(., aes(long, lat, color=distance)) +
  geom_segment(aes(xend = long + delta_long, yend = lat + delta_lat), # add a vector plot
               arrow = arrow(length = unit(0.1,"cm")),lwd=2) +
  coord_equal()+ # fix the coords
  # borders("usa")+
  scale_x_continuous(limits = c(-150,-120))+
  scale_color_viridis_c()+
  theme_dark()
```

<br>

## Plotting hurricane tracks ------------------------------------------------------------------
```{r dplyr::storms, cache=T}
library(tidyverse)
dplyr::storms
names(storms)

# bad!
storms %>%
  ggplot(data=., aes(long,lat,size=category))+
  geom_point()+
  # borders("world")+
  # scale_x_continuous(limits=c(-120,0))+
  # scale_y_continuous(limits = c(-10,55))+
  borders('world', fill = "darkgreen")

# still bad!
storms %>% 
  arrange(wind) %>% 
  ggplot(data=., aes(long,lat,size=category,color=wind))+
  # borders("coast")+
  geom_point()+
  scale_color_viridis_c()+
  scale_x_continuous(limits=c(-130,0))+
  scale_y_continuous(limits = c(-10,55))


```

## A better way, but it's a bit tricky
```{r cache=T, message=F, warning=F}

library(sf); library(fasterize);
dplyr::storms
x <- map('world', plot=F, fill=T) %>% st_as_sf()
blank_slate <- raster::raster(x, res=1)
r <- fasterize::fasterize(x, blank_slate)
rat_world <- r %>% raster::as.data.frame(., xy=T) %>% 
  rename(long=x, lat=y) %>% as_tibble()

ss_rat_world <- rat_world %>% 
  filter(lat > -15 & lat < 60 & long > -150 & long < 0)

ss_rat_world %>% 
  filter(is.na(layer)==F) %>% 
  ggplot(data=., aes(long,lat))+geom_raster()+
  geom_point(data=storms, aes(x=long,y=lat,color=wind,size=category),alpha=0.25)+
  scale_color_viridis_c("B", begin=0.2)+
  coord_equal()+
  theme_bw()

```

## Time series of hurricanes
```{r cache=T, message=F, warning=F}
library(tidyverse); library(lubridate)
dplyr::storms
storms %>% group_by(year) %>% summarize(nobs=n()) %>% ggplot(data=., aes(year,nobs))+geom_line()


storms %>% 
  group_by(month) %>% 
  summarize(wind_25=quantile(wind,0.025), 
            wind_50=median(wind), 
           wind_75=quantile(wind, 0.95)) %>% 
  ungroup() %>% 
  ggplot(data=., aes(month, wind_50))+
  geom_ribbon(aes(x=month, ymax=wind_75, ymin=wind_25),lty=0,alpha=0.25)+
  geom_line()+
  labs(x="Month",y="Wind speed [knots]",title = "95% quantile rate of hurricane wind speed")

# swap 'month' for 'year'
storms %>% group_by(year) %>% summarize(wind_25=quantile(wind,0.025), 
                                         wind_50=median(wind), 
                                         wind_75=quantile(wind, 0.95)) %>% 
  ungroup() %>% 
  ggplot(data=., aes(year, wind_50))+
  geom_ribbon(aes(x=year, ymax=wind_75, ymin=wind_25),lty=0,alpha=0.25)+
  geom_line()+
  labs(x="Month",y="Wind speed [knots]",title = "95% quantile rate of hurricane wind speed")

# advanced!
p = c(0.025, 0.25,0.5,0.75,0.975)
storms %>% 
  group_by(month) %>% 
  summarise(quantiles = list(sprintf("%1.0f%%", p*100)),
            wind = list(quantile(wind, p))) %>% 
  unnest %>% 
  ggplot(data=., aes(month, wind, color=quantiles))+
  geom_line()+
  scale_color_viridis_d(end=0.8)

```


<br>
<br>

## PCA example with columns scaling -------------------------------------------------------------
```{r message=FALSE, warning=FALSE, cache=F, paged.print=TRUE}
library(tidyverse); 
rm(iris); data('iris'); # restoring default version of iris dataset

iris %>%
  prcomp(~.-Species, data=.) %>% # run principle components on all vars, except Species
  biplot(main='BAD!')                       # plot biplot

iris %>% 
  group_by(Species) %>%          # grouping var
  mutate_all(scale) %>%          # center vars over zero, and divide by sd
  ungroup() %>% 
  prcomp(~.-Species, data=.) %>% # run principle components on all vars, except Species
  biplot(main='Good')                       # plot biplot

```

<br>
<br>
<br>

# Super advanced dplyr ----------------------------------------------------
## inspired by: [Suzan Baert](https://twitter.com/SuzanBaert) and modified from her github [repo]( https://github.com/suzanbaert/RLadies_RoCur/blob/master/Dplyr_tricks.pdf) tutorial on advanced dplyr
```{r cache=T}
# using !! "bang"
vars <- c("lat","long","wind")
storms %>% select(!!vars)

# select columns by regex
who %>% names # lots of column names
who %>% select(country, year, matches("*2534")) # select country, year, and columns with '2534' in the name

# rename columns with regex
library(stringr);
iris %>% 
  as_tibble() %>% 
  rename_all(tolower) %>% 
  rename_all(~str_replace_all(., "\\.","_"))

# mutate *observation* names
storms %>% 
  select(name,year,status) %>% 
  mutate_all(tolower) %>% # Amy -> amy
  mutate_all(~str_replace_all(., " ","_")) # 'tropical depression' -> 'tropical_depression'

# find highest values
storms %>% 
  top_n(5, wind) # storms with 5 highest windspeeds

# making new vars from conditions
starwars %>%
  select(name, species, homeworld, birth_year, hair_color) %>%
  mutate(new_group = case_when(
    species == "Droid" ~ "Robot",
    homeworld == "Tatooine" & hair_color == "blond" ~ "Blond Tatooinian",
    homeworld == "Tatooine" ~ "Other Tatooinian",
    hair_color == "blond" ~ "Blond non-Tatooinian",
    TRUE ~ "Other Human"))
```


<br>
<br>
<br>


# (6) SPATIAL METHODS + TIDYVERSE ---------------------------------------------
### The sp library is commonly used for spatial vector data, although it is not very compatible with tidyverse methods
### The developing 'simple features' (sf) library aims to be a tidyverse compliant sp alternative
```{r cache=T}
# Required libraries: 
library(tidyverse); 
library(lubridate); 
library(sf); 

dir.create("data/SouthAmerica")
unzip(zipfile = "data/SouthAmerica.zip", exdir = "data/SouthAmerica")
list.files('data/SouthAmerica/')

SA <- sf::st_read("data/SouthAmerica/SouthAmerica.shp")

plot(SA) # blah, not ideal
plot(SA["SQKM"]) # base R method - a little better, but not so easy to control

# Way better with ggplot
ggplot() + 
  geom_sf(data=SA, aes(fill=SQKM))+
  scale_fill_viridis_c()

ggplot() + geom_sf(data=SA, aes(fill=log(SQKM, base = 10)))+
  scale_fill_viridis_c("Area [log10(km2)]")

# SA %>% mutate(population=ifelse(POP2007>0, POP2007,1)) %>% 
#   select(population) %>% pull(population)
#   ggplot()+
#   geom_sf(data=SA, aes(fill=population))+
#   scale_fill_viridis_c()



# SA %>% 
#   ggplot(data=., aes())+geom_sf(fill="SQKM")+
#   geom_point(data=data.frame(lat=0,lon=-80),aes(lat,lon),col='red')
# 
# ggplot(data=SA, aes())+geom_sf()+
#   geom_point(data=data.frame(lat=0,lon=-80),aes(lat,lon),col='red')


```

<br>
<br>

## Joining multiple data sources with spatial features
```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE, cache=T}
# get lifeExp from gapminder
dat <- gapminder::gapminder
dat$year %>% summary
pop2000s <- dat %>% 
  filter(year>=2000) %>% 
  group_by(country) %>% 
  summarize(le=mean(lifeExp, na.rm=T))

library(maps)
world = st_as_sf(map('world', plot = FALSE, fill = TRUE))
laea = st_crs("+proj=laea +lat_0=30 +lon_0=-95") # Lambert equal area
new_world <- st_transform(world, laea) # spatial transform of projection
g = st_graticule(world) # the lines on a map
plot(st_geometry(g), axes = TRUE)

plot(world, graticule = TRUE, key.pos = NULL, axes = TRUE)

# are the countries in the gapminder data in the world map?
(unique(dat$country) %in% world$ID) %>% table()

# join the gapminder data with the spatial data
tmp2 <- left_join(world, 
                  pop2000s %>% rename(ID=country), # needed to match by 'ID' 
                  by="ID") 

ggplot()+                           # initiate ggplot
  geom_sf(data=tmp2, aes(fill=le))+ # add spatial data 
  scale_fill_viridis_c("Life Expectancy [yrs]", option="B")+  # use the inferno color palette
  labs(x="Longitude",y="Latitude", title = "Life Expectancy - 2000s", subtitle = "source: gapminder")
```

<br>
<br>

## Plotting raster data with **ggplot2**
This works for visualizing single bands of smallish rasters (< 1000x1000) 
```{r cache=T}
library(tidyverse)
# Calculate NDVI from Landsat 7 -------------------------------------------
tif = system.file("tif/L7_ETMs.tif", package = "stars")
(r = raster::stack(tif))
raster::plot(r) # Not ideal.

l7 <- raster::as.data.frame(r, xy=T) %>% as_tibble()
l7 <- l7 %>%
  mutate(ndvi=(L7_ETMs.4-L7_ETMs.3)/(L7_ETMs.4+L7_ETMs.3))
l7 %>% 
  ggplot(data=., aes(x,y,fill=ndvi))+
  geom_raster()+
  coord_equal()+
  theme_bw()+
  scale_fill_viridis_c("NDVI")+
  labs(x="UTM X [m]",y="UTM Y [m]")+
  theme(legend.position = c(0.9,0.15),
        legend.title = element_text(size=15, face = 'bold'),
        legend.text = element_text(size=10),
        axis.title.x = element_text(size=25), 
        axis.text.x = element_text(size=15), 
        axis.title.y = element_text(size=25), 
        axis.text.y = element_text(size=15))

```

<br>
<br>
<br>

# A peak into stars! ------------------------------------------------------------------
# stars is developing package for dealing with spatial raster and vector data
# It's tidyverse compliant, and is/will be much better suited for processing large spatial data in R
# [stars library](https://r-spatial.github.io/stars/)

# **CAUTIONARY NOTE!** _If_ you are processing Gbs worth of raster or other spatiotemporal data, consider doing it in Python.
```{r cache=T}
library(stars)
tif = system.file("tif/L7_ETMs.tif", package = "stars")
(r = raster::stack(tif))
raster::plot(r) # Not ideal.

(x = read_stars(tif))
plot(x) # much improved (see ?plot.stars)
x[,,,1] %>% plot # plot band 1
x[,,,2] %>% plot # plot band 2
x[,1:100,1:100,1] %>% plot # plot spatial subset
x[,1:10,1:10,c(1,2,3)] %>% plot
```

## Plot an RGB with bands 5,4,3
```{r cache=T, message=F, warning=F}
library(stars)
tif = system.file("tif/L7_ETMs.tif", package = "stars")

(x = read_stars(tif))
image(x, rgb=c(5,4,3), axes=T)
```
